# Results

```{r packages}
  library(rgdal)
  library(dplyr)
  library(leaflet)
  library(readxl)
  library(leaflet.extras)
```

## Choropleth Map showing commuter-packed counties and covid cases.

In our exploratory analysis, we discovered that commuters have a positive correlation with the Covid case grow. The following choropleth map shows some interesting patterns:

```{r choropleth}
nys_spdf <- readOGR( 
  dsn= paste0(getwd(),"/dataset/Map/NYS") , 
  layer="cty036",
  verbose=FALSE
)
# Loading Geographic Data
nys_geo <- nys_spdf@data[,c('NAME', 'POP2000')]
lat = c()
lng = c()
for (i in c(1:67)){
  coord = nys_spdf@polygons[[i]]@labpt
  lat <- append(lat, coord[2])
  lng <- append(lng, coord[1])
}
nys_geo$lati <- lat
nys_geo$long <- lng
nys_geo <- nys_geo %>% group_by(NAME) %>%
  summarize(lati = mean(lati, na.rm=TRUE), long = mean(long, na.rm=TRUE))

# Loading Covid Case Data
nys_case <- read.csv("./dataset/Covid/NYS_Covid_Case.csv",
                     , sep=",", dec=".")

# Loading Work_Residence Data
commute <-
  read_excel("./dataset/Work_Residence/Work_Residence.xlsx")  
commute$commuter <- commute$Worked_outside_PoR / commute$Total_Labor

info<-merge(x=nys_case, y=commute, by.x="County", by.y = "Geographic", all= T)
info<-merge(x=info, y=nys_geo, by.x="County", by.y = "NAME", all= T)
info <- as_tibble(info)

order <- nys_spdf@data$NAME
nys_spdf@data <- merge(x=nys_spdf@data, y=commute,
                      by.x="NAME", by.y = "Geographic",
                      all= T, sort=FALSE)

nys_spdf@data <- nys_spdf@data %>%
  slice(match(order, NAME))

info$TestDate <- as.Date(as.character(info$TestDate), format = "%m/%d/%y")

info$sqtcum <- sqrt(info$Cumulative_Positives)
info$logcum <- log(info$Cumulative_Positives+1)
info$sqtnew <- sqrt(info$New_Positives)
info$lognew <- log(info$New_Positives+1)

bin_max = max(info$commuter)
bin_min = min(info$commuter)

info_final <- info[info$TestDate == '2020-11-28',]
info_final <- as_tibble(info_final)

palette <- colorBin(c('#fee0d2',  
                      '#fcbba1',
                      '#fc9272',
                      '#fb6a4a',
                      '#ef3b2c',
                      '#cb181d',
                      '#a50f15',
                      '#67000d'), 
                    bins = c(bin_min, bin_max/8, 2*bin_max/8,
                             3*bin_max/8, 4*bin_max/8,
                             5*bin_max/8, 6*bin_max/8,
                             7*bin_max/8, bin_max))

info_mar <- info[info$TestDate <= '2020-03-28',]
monthly <- info_mar %>% group_by(County) %>% summarize(monthly = sum(New_Positives))
info_mar <-merge(x=info_mar, y=monthly, by.x="County", by.y = "County", all= T)
info_mar <- as_tibble(info_mar)

info_apr <- info[(info$TestDate > '2020-03-28')&(info$TestDate <= '2020-04-28'),]
monthly <- info_apr %>% group_by(County) %>% summarize(monthly = sum(New_Positives))
info_apr <-merge(x=info_apr, y=monthly, by.x="County", by.y = "County", all= T)
info_apr <- as_tibble(info_apr)

info_may <- info[(info$TestDate > '2020-04-28')&(info$TestDate <= '2020-05-28'),]
monthly <- info_may %>% group_by(County) %>% summarize(monthly = sum(New_Positives))
info_may <-merge(x=info_may, y=monthly, by.x="County", by.y = "County", all= T)
info_may <- as_tibble(info_may)

info_jun <- info[(info$TestDate > '2020-05-28')&(info$TestDate <= '2020-06-28'),]
monthly <- info_jun %>% group_by(County) %>% summarize(monthly = sum(New_Positives))
info_jun <-merge(x=info_jun, y=monthly, by.x="County", by.y = "County", all= T)
info_jun <- as_tibble(info_jun)

info_jul <- info[(info$TestDate > '2020-06-28')&(info$TestDate <= '2020-07-28'),]
monthly <- info_jul %>% group_by(County) %>% summarize(monthly = sum(New_Positives))
info_jul <-merge(x=info_jul, y=monthly, by.x="County", by.y = "County", all= T)
info_jul <- as_tibble(info_jul)

info_aug <- info[(info$TestDate > '2020-07-28')&(info$TestDate <= '2020-08-28'),]
monthly <- info_aug %>% group_by(County) %>% summarize(monthly = sum(New_Positives))
info_aug <-merge(x=info_aug, y=monthly, by.x="County", by.y = "County", all= T)
info_aug <- as_tibble(info_aug)

info_sep <- info[(info$TestDate > '2020-08-28')&(info$TestDate <= '2020-09-28'),]
monthly <- info_sep %>% group_by(County) %>% summarize(monthly = sum(New_Positives))
info_sep <-merge(x=info_sep, y=monthly, by.x="County", by.y = "County", all= T)
info_sep <- as_tibble(info_sep)

info_oct <- info[(info$TestDate > '2020-09-28')&(info$TestDate <= '2020-10-28'),]
monthly <- info_oct %>% group_by(County) %>% summarize(monthly = sum(New_Positives))
info_oct <-merge(x=info_oct, y=monthly, by.x="County", by.y = "County", all= T)
info_oct <- as_tibble(info_oct)

info_nov <- info[(info$TestDate > '2020-10-28')&(info$TestDate <= '2020-11-28'),]
monthly <- info_nov %>% group_by(County) %>% summarize(monthly = sum(New_Positives))
info_nov <-merge(x=info_nov, y=monthly, by.x="County", by.y = "County", all= T)
info_nov <- as_tibble(info_nov)

# Construct Leaflet Map
mymap <- leaflet() %>% 
  addProviderTiles("Esri.WorldGrayCanvas",
                   options = tileOptions(minZoom=2, maxZoom=10)
  )%>%
  addPolygons(data = nys_spdf, 
              fillColor = ~palette(nys_spdf@data$commuter),
              fillOpacity = 0.6,         ## how transparent do you want the polygon to be?
              color = "darkgrey",       ## color of borders between districts
              weight = 1.5,            ## width of borders
              group="background_group",
              label = nys_spdf@data$NAME,
              popup = ~ paste0("<b>", nys_spdf@data$NAME, 
                               "</b><br>Total Labor: <br>", nys_spdf@data$Total_Labor,
                               "</b><br>Outside PoR: <br>", nys_spdf@data$Worked_outside_PoR)
  )%>%
  addCircleMarkers(
    group = "2020-Mar",
    data = info_mar,
    lng = ~long,
    lat = ~lati, 
    radius = ~1.3*log(monthly+1)+1, 
    color = "Teal",
    fillOpacity=0.1,
    stroke = FALSE, 
    popup = ~ paste0("<b>", County, "</b><br>Monthly Confirmed Cases: <br>", monthly)
  )%>%
  addCircleMarkers(
    group = "2020-Apr",
    data = info_apr,
    lng = ~long,
    lat = ~lati, 
    radius = ~1.3*log(monthly+1)+1, 
    color = "Teal", 
    opacity = 0.6,
    stroke = FALSE, 
    popup = ~ paste0("<b>", County, "</b><br>Monthly Confirmed Cases: <br>", monthly)
  )%>%
  addCircleMarkers(
    group = "2020-May",
    data = info_may,
    lng = ~long,
    lat = ~lati, 
    radius = ~1.3*log(monthly+1)+1, 
    color = "Teal", 
    opacity = 0.6,
    stroke = FALSE, 
    popup = ~ paste0("<b>", County, "</b><br>Monthly Confirmed Cases: <br>", monthly)
  )%>%
  addCircleMarkers(
    group = "2020-Jun",
    data = info_jun,
    lng = ~long,
    lat = ~lati, 
    radius = ~1.3*log(monthly+1)+1, 
    color = "Teal", 
    opacity = 0.6,
    stroke = FALSE, 
    popup = ~ paste0("<b>", County, "</b><br>Monthly Confirmed Cases: <br>", monthly)
  )%>%
  addCircleMarkers(
    group = "2020-Jul",
    data = info_jul,
    lng = ~long,
    lat = ~lati, 
    radius = ~1.3*log(monthly+1)+1, 
    color = "Teal",
    opacity = 0.6,
    stroke = FALSE, 
    popup = ~ paste0("<b>", County, "</b><br>Monthly Confirmed Cases: <br>", monthly)
  )%>%
  addCircleMarkers(
    group = "2020-Aug",
    data = info_aug,
    lng = ~long,
    lat = ~lati, 
    radius = ~1.3*log(monthly+1)+1, 
    color = "Teal", 
    opacity = 0.6,
    stroke = FALSE, 
    popup = ~ paste0("<b>", County, "</b><br>Monthly Confirmed Cases: <br>", monthly)
  )%>%
  addCircleMarkers(
    group = "2020-Sep",
    data = info_sep,
    lng = ~long,
    lat = ~lati, 
    radius = ~1.3*log(monthly+1)+1, 
    color = "Teal",
    opacity = 0.6,
    stroke = FALSE, 
    popup = ~ paste0("<b>", County, "</b><br>Monthly Confirmed Cases: <br>", monthly)
  )%>%
  addCircleMarkers(
    group = "2020-Oct",
    data = info_oct,
    lng = ~long,
    lat = ~lati, 
    radius = ~1.3*log(monthly+1)+1, 
    color = "Teal", 
    opacity = 0.6,
    stroke = FALSE, 
    popup = ~ paste0("<b>", County, "</b><br>Monthly Confirmed Cases: <br>", monthly)
  )%>%
  addCircleMarkers(
    group = "2020-Nov",
    data = info_nov,
    lng = ~long,
    lat = ~lati, 
    radius = ~1.3*log(monthly+1)+1, 
    color = "Teal",
    opacity = 0.6,
    stroke = FALSE, 
    popup = ~ paste0("<b>", County, "</b><br>Monthly Confirmed Cases: <br>", monthly)
  )%>%
  addLayersControl(
    baseGroups = c("2020-Mar", "2020-Apr", "2020-May", "2020-Jun", "2020-Jul", "2020-Aug",
                   "2020-Sep", "2020-Oct", "2020-Nov"), 
    position = "bottomleft", 
    options = layersControlOptions(collapsed = FALSE)
  )%>% 
  addLegend(position = 'topright', 
            colors = c('#fee0d2',
                       '#fcbba1',
                       '#fc9272',
                       '#fb6a4a',
                       '#ef3b2c',
                       '#cb181d',
                       '#a50f15',
                       '#67000d'), 
            labels = c(paste(trunc(100*bin_min), '%', sep=""),"","","",
                       "","","",paste(trunc(100*bin_max), '%', sep="")),  
            opacity = 0.6,      
            title = "Labor working <br>outside county<br> of residence"
  )%>%
  addLegend(position = 'topleft',
            colors = 'Teal',
            labels = "New Postive <br>During the Month"
  )%>%
  addResetMapButton()
mymap  
```

The heatmap shows the percentage of labor working outside county of residence. Deeper the color of the county is, the more ratio of workforce commutes outside the county. The teal colored circles represent the number of cumulative cases within each month of 2020. The general pattern indicates and deeper the color is, the more likely that corresponding counties and the surrounding counties experience sharp growth in covid numbers.